---
title: "Thesis Analysis"
author: "Rhea Sarma"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    toc: true
    toc_depth: '4'
  pdf_document:
    toc: true
    toc_depth: '4'
header-includes:
- \usepackage{graphicx}
- \usepackage{float}
- \setlength{\emergencystretch}{3em}
geometry: a4paper, margin=1in
fontsize: 11pt
editor_options:
  chunk_output_type: console
---

Checking the relationship between BIS, BAS, meal intake and EAH intake variables.

```{r}
library(haven)
library(dplyr)
library(ggplot2)
```

1.  Load dataset

```{r}
thesis_data<-read.csv("~/Desktop/Rhea MS thesis/MS_thesis/data/thesis_data.csv")
```

2.  My IV's of interest are `bis` `bas` `bas_funseeking` `bas_drive` `bas_rewardresp`

My DV's of interest are `meal_grams_consumed` `meal_kcal_consumed` `eah_grams_consumed_foodonly` `eah_kcal_consumed`

1.  Checking normality and homogeneity of variance assumptions and conducting visualizations

```{r}
# For variable BIS
qqnorm(thesis_data$bis)
qqline(thesis_data$bis)
shapiro.test(thesis_data$bis) #met normality


#Visualizing BIS data by Study, boxplot and histogram
ggplot(thesis_data, aes(x=Study, y=bis)) + geom_boxplot() + geom_jitter(height = NULL) #jitters add all the data points
ggplot(thesis_data,aes(bis))+ geom_histogram(bins=15)+facet_grid(.~Study)

# For variable BAS
qqnorm(thesis_data$bas)
qqline(thesis_data$bas)
shapiro.test(thesis_data$bas) #met normality


#Visualizing BAS data by Study, boxplot and histogram
ggplot(thesis_data, aes(x=Study, y=bas)) + geom_boxplot() + geom_jitter(height = NULL) 
ggplot(thesis_data,aes(bas))+geom_histogram(bins=15)+facet_grid(.~Study)

# For variable BAS funseeking
qqnorm(thesis_data$bas_funseeking)
qqline(thesis_data$bas_funseeking)
shapiro.test(thesis_data$bas_funseeking) #not normal


#Visualizing BAS funseeking data by Study, boxplot and histogram
ggplot(thesis_data, aes(x=Study, y=bas_funseeking)) +geom_boxplot() +geom_jitter(height = NULL) 
ggplot(thesis_data,aes(bas_funseeking))+geom_histogram(bins=15)+facet_grid(.~Study)

# For variable BAS drive
qqnorm(thesis_data$bas_drive)
qqline(thesis_data$bas_drive)
shapiro.test(thesis_data$bas_drive) #not normal

#Visualizing BAS drive data by Study, boxplot and histogram
ggplot(thesis_data, aes(x=Study, y=bas_drive)) + geom_boxplot() + geom_jitter(height = NULL) 
ggplot(thesis_data,aes(bas_drive))+geom_histogram(bins=15)+ facet_grid(.~Study)

# For variable BAS reward responsive
qqnorm(thesis_data$bas_rewardresp)
qqline(thesis_data$bas_rewardresp)
shapiro.test(thesis_data$bas_rewardresp) #not normal

#Visualizing BAS reward responsive data by Study, boxplot and histogram
ggplot(thesis_data, aes(x=Study, y=bas_rewardresp)) +geom_boxplot() + geom_jitter(height = NULL) 
ggplot(thesis_data,aes(bas_rewardresp))+geom_histogram(bins=15)+ facet_grid(.~Study)


# For variable Meal consumed in grams
qqnorm(thesis_data$meal_grams_consumed)
qqline(thesis_data$meal_grams_consumed)
shapiro.test(thesis_data$meal_grams_consumed) #not normal

#Visualizing Meal consumed in grams data by Study, boxplot and histogram
ggplot(thesis_data, aes(x=Study, y=meal_grams_consumed)) + geom_boxplot() + geom_jitter(height = NULL) 
ggplot(thesis_data,aes(meal_grams_consumed))+geom_histogram(bins=15)+facet_grid(.~Study)


# For variable Meal consumed in kcal
qqnorm(thesis_data$meal_kcal_consumed)
qqline(thesis_data$meal_kcal_consumed)
shapiro.test(thesis_data$meal_kcal_consumed) #not normal

#Visualizing Meal consumed in kcal data by Study, boxplot and histogram
ggplot(thesis_data, aes(x=Study, y=meal_kcal_consumed)) + geom_boxplot() + geom_jitter(height = NULL) 
ggplot(thesis_data,aes(meal_kcal_consumed))+geom_histogram(bins=15)+facet_grid(.~Study)


# For variable EAH consumed in grams
qqnorm(thesis_data$eah_grams_consumed_foodonly)
qqline(thesis_data$eah_grams_consumed_foodonly)
shapiro.test(thesis_data$eah_grams_consumed_foodonly) #not normal

#Visualizing EAH consumed in grams data by Study, boxplot and histogram
ggplot(thesis_data, aes(x=Study, y=eah_grams_consumed_foodonly)) +geom_boxplot() + geom_jitter(height = NULL) 
ggplot(thesis_data,aes(eah_grams_consumed_foodonly))+geom_histogram(bins=15)+facet_grid(.~Study)


# For variable EAH consumed in kcal
qqnorm(thesis_data$eah_kcal_consumed)
qqline(thesis_data$eah_kcal_consumed)
shapiro.test(thesis_data$eah_kcal_consumed) #not normal

#Visualizing EAH consumed in kcal data by Study, boxplot and histogram
ggplot(thesis_data, aes(x=Study, y=eah_kcal_consumed)) + geom_boxplot() + geom_jitter(height = NULL) 
ggplot(thesis_data,aes(eah_kcal_consumed))+geom_histogram(bins=15)+facet_grid(.~Study)

```

While some of the variables didn't pass the shapiro normality test but after looking at the qqplots of these variables we can consider them as normal.

3.Corrplots between the interested variables

```{r}
library(corrplot)

cor_all <- cor(thesis_data[c("bis", "bas","bas_funseeking","bas_drive", "bas_rewardresp", "pre_eah_fullness", "meal_grams_consumed",
                             "meal_kcal_consumed","eah_kcal_consumed", "eah_grams_consumed_foodonly")], use = "pairwise.complete.obs")

cor_reach <- cor(thesis_data[thesis_data$Study == "REACH", c("bis", "bas","bas_funseeking","bas_drive", "bas_rewardresp", "pre_eah_fullness", "meal_grams_consumed","meal_kcal_consumed","eah_kcal_consumed", "eah_grams_consumed_foodonly")], use = "pairwise.complete.obs")

cor_fb <- cor(thesis_data[thesis_data$Study == "FOOD_BRAIN", c("bis", "bas","bas_funseeking","bas_drive", "bas_rewardresp", "pre_eah_fullness", "meal_grams_consumed","meal_kcal_consumed","eah_kcal_consumed", "eah_grams_consumed_foodonly")], use = "pairwise.complete.obs")


corrplot(cor_all, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

corrplot(cor_reach, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

corrplot(cor_fb, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

\pagebreak

**AIM 1 : Does BIS and BAS influences eating in absence of hunger?**

```{r}
# 1. Relationship between BIS and EAH

## BIS is not associated with EAH gram intake
cor.test(thesis_data$bis, thesis_data$eah_grams_consumed_foodonly,
         use = "pairwise.complete.obs") 

## BIS is associated with EAH kcal intake
cor.test(thesis_data$bis, thesis_data$eah_kcal_consumed, 
         use = "pairwise.complete.obs")

#2. Relationship between BAS and EAH

## BAS is not associated with EAH gram intake 
cor.test(thesis_data$bis, thesis_data$eah_grams_consumed_foodonly, 
         use = "pairwise.complete.obs") 

## BAS is not associated with EAH kcal intake
cor.test(thesis_data$bas, thesis_data$eah_kcal_consumed, 
         use = "pairwise.complete.obs")

#3. Relationship between BAS funseeking and EAH

## BAS funseeking is not associated with EAH gram intake 
cor.test(thesis_data$bas_funseeking, 
         thesis_data$eah_grams_consumed_foodonly, use = "pairwise.complete.obs") 
 
## BAS funseeking is not associated with EAH kcal intake
cor.test(thesis_data$bas_funseeking, 
         thesis_data$eah_kcal_consumed, use = "pairwise.complete.obs")

#4. Relationship between BAS drive and EAH

## BAS drive is not associated with EAH gram intake
cor.test(thesis_data$bas_drive,
         thesis_data$eah_grams_consumed_foodonly, use = "pairwise.complete.obs") 

## BAS drive is not associated with EAH kcal intake
cor.test(thesis_data$bas_drive,
         thesis_data$eah_kcal_consumed, use = "pairwise.complete.obs")

#5. Relationship between BAS reward responsive and EAH

## BAS reward responsive is not associated with EAH gram intake 
cor.test(thesis_data$bas_rewardresp, 
         thesis_data$eah_grams_consumed_foodonly, use = "pairwise.complete.obs") 

## BAS reward responsive is not associated with EAH kcal intake
cor.test(thesis_data$bas_rewardresp, 
         thesis_data$eah_kcal_consumed, use = "pairwise.complete.obs")

```

Based on the correlations and corr plots we can see that `bis` is negativley related with `eah_kcal_consumed`

4.  Conducting levene's test on `bis` and `eah_kcal_consumed` by Study and then t-test on `bis` and `eah_kcal_consumed` to see if there is study effect.

```{r}
library(car)
#For bis
leveneTest(
bis~as.factor(Study), data=thesis_data)

#For eah_kcal_consumed
leveneTest(
eah_kcal_consumed~as.factor(Study), data=thesis_data)

#Assumptions for homegenity in variances are met

#Conducting indepedent t-test on bis to see study effect
t.test(thesis_data[thesis_data$Study == "REACH",]$bis, 
       thesis_data[thesis_data$Study  == "FOOD_BRAIN",]$bis,
       alternative = "two.sided", var.equal = TRUE)

#Conducting indepedent t-test on bis to see study effect
t.test(thesis_data[thesis_data$Study == "REACH",]$eah_kcal_consumed, thesis_data[thesis_data$Study  == "FOOD_BRAIN",]$eah_kcal_consumed, 
       alternative = "two.sided", var.equal = TRUE)


#Visualizing 
ggplot(thesis_data, aes(x=Study, y=bis)) + 
geom_boxplot() +
geom_jitter(height = NULL)

ggplot(thesis_data, aes(x=Study, y=eah_kcal_consumed)) + 
  geom_boxplot() + 
  geom_jitter(height = NULL) 
thesis_data$Age.in.years

```

Based on these results, BIS scores are higher in REACH study and EAH consumption in kcal was similar in both the studies.

5.  Linear regressions predicting EAH with covariates [sex + child bmi + age in years+ pre eah fullness, income] (no interactions) -- no effects

```{r}
model1 <- lm(eah_kcal_consumed ~ bis + Sex + Age.in.years + Child.BMI 
             + Income + pre_eah_fullness, data = thesis_data)
summary(model1)
```

Based on the summary of the model of this multiple regression model we can see that higher BIS scores predicts a lower food intake in kcal ( B=-60, p=0.04) in EAH paradigm.

\pagebreak

**AIM 2: Does maternal risk status moderates the relationship between BIS, BAS and EAH**

```{r}
#Conducting 2 way ANOVA to see the effect of maternal risk status on EAH in kcal across both the studies

thesis_data$Maternal.risk.status <- as.factor(thesis_data$Maternal.risk.status)
thesis_data$Study <- as.factor(thesis_data$Study)

anova_model <- aov(eah_kcal_consumed ~ Maternal.risk.status * Study, data = thesis_data)
summary(anova_model)

#Visualization
maternal_data <- thesis_data %>% filter(!is.na(Study), !is.na(Maternal.risk.status))
library(ggplot2)
ggplot(maternal_data, aes(x = Study, y = eah_kcal_consumed, fill = Maternal.risk.status)) +
  geom_boxplot() +
  labs(title = "Effect of Maternal Risk Status on EAH Intake Across Studies",
       x = "Study",
       y = "EAH Intake (kcal)",
       fill = "Maternal Risk Status")


```

There wasn’t any main effect of maternal risk status ( p=0.87) and study(p=0.89) on EAH intake. Moreover, there wasn’t any interaction effect between maternal risk status and study on EAH intake (p=0.87). Hence we can conclude that effect of maternal risk status on EAH paradigm was consistent in both the studies.

```{r}

#Conducting moderation analysis adjusting for covariates
lm_eah_kcal_risk <- lm(eah_kcal_consumed ~ bis*Maternal.risk.status + pre_eah_fullness + Study + Child.BMI + Age.in.years + Sex + Income, data = thesis_data)

summary(lm_eah_kcal_risk)

```

After adjusting for sex, child age, child bmi, parent income, Pre EAH fullness scores and study, we conducted moderation analysis. The results revealed a significant interaction between BIS_total and maternal risk status but only for children with low-risk of obesity (B= -131.1, p=0.02), indicating that children at low risk for obesity, a 1 unit increase in BIS corresponds to 131 less kcal consumed during EAH , while children at high risk for obesity, the association between BIS and EAH is not significant.

Making plots

```{r}
# remove rows with missing values for model variables -- will make it easier to save predicted values to model dataset
MISSING <- is.na(thesis_data$eah_kcal_consumed) |
           is.na(thesis_data$bis) |
           is.na(thesis_data$pre_eah_fullness) |
           is.na(thesis_data$Study) |
           is.na(thesis_data$Child.BMI) |
           is.na(thesis_data$Age.in.years) | 
          is.na(thesis_data$Sex) | is.na(thesis_data$Maternal.risk.status)

thesis_data_no_na <- subset(thesis_data, 
                      subset = !MISSING)

#Visualizations
# plot kcal (raw values)
ggplot(thesis_data, aes(x = bis, y = eah_kcal_consumed, color = factor(Maternal.risk.status))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, aes(group = Maternal.risk.status)) +
  labs(x = "bis", y = "eah_kcal_consumed", color = "Maternal.risk.status") + ylim(0, 900)

#3
ggplot(thesis_data_no_na, aes(x = bis, y = eah_kcal_consumed, color = factor(Maternal.risk.status))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, aes(group = Maternal.risk.status)) +
  labs(x = "bis", y = "eah_kcal_consumed (adjusted)", color = "Maternal.risk.status")
```

Adding and merging food item columns from both the studies

```{r}
#EAH sweet data for all in grams
thesis_data<-thesis_data %>%
    mutate(all_eah_brownies_g = coalesce(fb_eah_consumed_brownies_g, reach_eah_brownie_grams_consumed),
           all_eah_hersheys_g = coalesce(fb_eah_consumed_hersheys_g, reach_eah_kiss_grams_consumed),
           all_eah_icecream_g = coalesce(fb_eah_consumed_icecream_g, reach_eah_ice_cream_grams_consumed),
           all_eah_oreos_g = coalesce(fb_eah_consumed_oreos_g, reach_eah_oreo_grams_consumed),
           all_eah_skittles_g = coalesce(fb_eah_consumed_skittles_g, reach_eah_skittle_grams_consumed),
           all_eah_starbursts_g = coalesce(fb_eah_consumed_starbursts_g, reach_eah_starburst_grams_consumed))

#EAH salty data for all in grams
thesis_data<-thesis_data %>%
    mutate(all_eah_cornchips_g = coalesce(fb_eah_consumed_cornchips_g, reach_eah_corn_chip_grams_consumed),
           all_eah_popcorn_g = coalesce(fb_eah_consumed_popcorn_g, reach_eah_popcorn_grams_consumed),
           all_eah_pretzels_g = coalesce(fb_eah_consumed_pretzels_g, reach_eah_pretzel_grams_consumed))

#EAH sweet data for all in kcal
thesis_data<-thesis_data %>%
    mutate(all_eah_brownies_kcal = coalesce(fb_eah_consumed_brownies_kcal, reach_eah_brownie_kcal_consumed),
           all_eah_hersheys_kcal = coalesce(fb_eah_consumed_hersheys_kcal, reach_eah_kiss_kcal_consumed),
           all_eah_icecream_kcal = coalesce(fb_eah_consumed_icecream_kcal, reach_eah_ice_cream_kcal_consumed),
           all_eah_oreos_kcal = coalesce(fb_eah_consumed_oreos_kcal, reach_eah_oreo_kcal_consumed),
           all_eah_skittles_kcal = coalesce(fb_eah_consumed_skittles_kcal, reach_eah_skittle_kcal_consumed),
           all_eah_starbursts_kcal = coalesce(fb_eah_consumed_starbursts_kcal, reach_eah_starburst_kcal_consumed))

#EAH salty data for all in kcal
thesis_data<-thesis_data %>%
    mutate(all_eah_cornchips_kcal = coalesce(fb_eah_consumed_cornchips_kcal, reach_eah_corn_chip_kcal_consumed),
           all_eah_popcorn_kcal = coalesce(fb_eah_consumed_popcorn_kcal, reach_eah_popcorn_kcal_consumed),
           all_eah_pretzels_kcal = coalesce(fb_eah_consumed_pretzels_kcal, reach_eah_pretzel_kcal_consumed))
```

Additional analysis on seperate food items

```{r}
library(corrplot)

cor_all_eah_sweet_grams <- cor(thesis_data[c("bis","bas","bas_funseeking","bas_drive", "bas_rewardresp","all_eah_brownies_g","all_eah_hersheys_g", "all_eah_icecream_g", "all_eah_oreos_g","all_eah_skittles_g", "all_eah_starbursts_g")], use = "pairwise.complete.obs")

cor_all_eah_salty_grams <- cor(thesis_data[c("bis","bas","bas_funseeking","bas_drive", "bas_rewardresp", "all_eah_cornchips_g", "all_eah_popcorn_g", "all_eah_pretzels_g")], use = "pairwise.complete.obs")

cor_all_eah_sweet_kcal <- cor(thesis_data[c("bis","bas","bas_funseeking","bas_drive", "bas_rewardresp","all_eah_brownies_kcal","all_eah_hersheys_kcal", "all_eah_icecream_kcal", "all_eah_oreos_kcal","all_eah_skittles_kcal", "all_eah_starbursts_kcal")], use = "pairwise.complete.obs")

cor_all_eah_salty_kcal <- cor(thesis_data[c("bis","bas","bas_funseeking","bas_drive", "bas_rewardresp", "all_eah_cornchips_kcal", "all_eah_popcorn_kcal", "all_eah_pretzels_kcal")], use = "pairwise.complete.obs")


corrplot(cor_all_eah_sweet_grams, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 100)

corrplot(cor_all_eah_salty_grams, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 100)

corrplot(cor_all_eah_sweet_kcal, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 100)

corrplot(cor_all_eah_salty_kcal, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 100)
```
