---
title: "Thesis analysis"
author: "Rhea Sarma" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
header-includes: 
  \usepackage{graphicx}
editor_options: 
  chunk_output_type: console
---

This document will have the demographic analysis part for the MS thesis.

1.  Load libraries

```{r}
library(haven)
library(dplyr)
library(tidyverse)
library(psych)
library(gtsummary)

```

2.  Load the cleaned and complied dataset

```{r}
getwd()
thesis_data<- read.csv("~/Desktop/Rhea MS thesis/MS_thesis/data/data_compiled.csv")
```

3.  Changing integer variables to numeric to maintain variable consistency

    ```{r}
    str(thesis_data)
    thesis_data$income<-as.numeric(thesis_data$income)
    thesis_data$pre_meal_fullness<-as.numeric(thesis_data$pre_meal_fullness)
    thesis_data$fb_meal_consumed_mustard_kcal<-as.numeric(thesis_data$fb_meal_consumed_mustard_kcal)
    thesis_data$reach_meal_water_kcal_consumed<-as.numeric(thesis_data$reach_meal_water_kcal_consumed)
    str(thesis_data)

    # Making the dataset presentable
    colnames(thesis_data)[colnames(thesis_data) == "parent1_sex"] <- "Measured Parent"
    colnames(thesis_data)[colnames(thesis_data) == "study"] <- "Study"
    thesis_data$Study<-toupper(thesis_data$Study)
    colnames(thesis_data)[colnames(thesis_data) == "age_yr"] <- "Age in years"
    colnames(thesis_data)[colnames(thesis_data) == "risk_status_maternal"] <- "Maternal risk status"
    colnames(thesis_data)[colnames(thesis_data) == "sex"] <- "Sex"
    colnames(thesis_data)[colnames(thesis_data) == "child_bmi"] <- "Child BMI"
    colnames(thesis_data)[colnames(thesis_data) == "child_bmi_z"] <- "Child BMI z-score"
    colnames(thesis_data)[colnames(thesis_data) == "child_bmi_p"] <- "Child BMI percentile"
    colnames(thesis_data)[colnames(thesis_data) == "ethnicity"] <- "Ethinicity"
    colnames(thesis_data)[colnames(thesis_data) == "race"] <- "Race"
    colnames(thesis_data)[colnames(thesis_data) == "income"] <- "Income"

    # replace factor values with labels that match food and brain 
    thesis_data$Income<-factor(thesis_data$Income,levels=0:5,labels = c("<$20,000","$20,000-$35,000","$36,000-$50,000","$51,000-$75,000","$76,000-$100,000",">$100,000"))

    #### Export data. This is the dataset we will be using for all the analysis ####
    write.csv(thesis_data,"~/Desktop/Rhea MS thesis/MS_thesis/data/thesis_data.csv", row.names = FALSE)
    ```

4.  Inspecting dataset

```{r}
head(thesis_data)
table(thesis_data$study)
```

5.  Generating demographics table. `Unknown` indicates missing data

```{r}
thesis_data %>% select (!c(participant_id,pre_meal_fullness,pre_eah_fullness, meal_grams_consumed, meal_kcal_consumed, eah_grams_consumed_foodonly,bis, bas,bas_funseeking, bas_drive, bas_rewardresp, eah_kcal_consumed,fb_meal_consumed_applesauce_kcal, fb_meal_consumed_applesauce_g, fb_meal_consumed_carrot_g, fb_meal_consumed_carrot_kcal, fb_meal_consumed_cheese_sndwch_g, fb_meal_consumed_cheese_sndwch_kcal, fb_meal_consumed_cookies_g, fb_meal_consumed_cookies_kcal, fb_meal_consumed_ham_sndwch_g, fb_meal_consumed_ham_sndwch_kcal, fb_meal_consumed_milk_g, fb_meal_consumed_milk_kcal, fb_meal_consumed_pbj_sndwch_g, fb_meal_consumed_pbj_sndwch_kcal, fb_meal_consumed_potatochip_g, fb_meal_consumed_potatochip_kcal, fb_meal_consumed_turkey_sndwch_g, fb_meal_consumed_turkey_sndwch_kcal, fb_meal_consumed_ketchup_g, fb_meal_consumed_ketchup_kcal, fb_meal_consumed_mayo_g, fb_meal_consumed_mayo_kcal, fb_meal_consumed_mustard_g, fb_meal_consumed_mustard_kcal,fb_eah_consumed_brownies_g, fb_eah_consumed_brownies_kcal, fb_eah_consumed_cornchips_g,fb_eah_consumed_cornchips_kcal,fb_eah_consumed_hersheys_g,fb_eah_consumed_hersheys_kcal,fb_eah_consumed_icecream_g,fb_eah_consumed_icecream_kcal,fb_eah_consumed_oreos_g, fb_eah_consumed_oreos_kcal, fb_eah_consumed_popcorn_g,fb_eah_consumed_popcorn_kcal, fb_eah_consumed_pretzels_g, fb_eah_consumed_pretzels_kcal, fb_eah_consumed_skittles_g, fb_eah_consumed_skittles_kcal, fb_eah_consumed_starbursts_g,fb_eah_consumed_starbursts_kcal, fb_eah_consumed_water_g, reach_eah_brownie_grams_consumed, reach_eah_brownie_kcal_consumed, reach_eah_corn_chip_grams_consumed, reach_eah_corn_chip_kcal_consumed, reach_eah_ice_cream_grams_consumed,reach_eah_ice_cream_kcal_consumed,reach_eah_kiss_grams_consumed,reach_eah_kiss_kcal_consumed,reach_eah_oreo_grams_consumed, reach_eah_oreo_kcal_consumed, reach_eah_popcorn_grams_consumed, reach_eah_popcorn_kcal_consumed, reach_eah_pretzel_grams_consumed, reach_eah_pretzel_kcal_consumed, reach_eah_skittle_grams_consumed, reach_eah_skittle_kcal_consumed, reach_eah_starburst_grams_consumed, reach_eah_starburst_kcal_consumed,reach_eah_water_eah_grams_consumed,reach_eah_water_eah_kcal_consumed, reach_meal_grilled_cheese_grams_consumed,reach_meal_grilled_cheese_kcal_consumed, reach_meal_carrot_grams_consumed, reach_meal_carrot_kcal_consumed, reach_meal_chips_grams_consumed, reach_meal_chips_kcal_consumed, reach_meal_fruit_grams_consumed, reach_meal_fruit_kcal_consumed, reach_meal_ketchup_grams_consumed, reach_meal_ketchup_kcal_consumed, reach_meal_ranch_grams_consumed, reach_meal_ranch_kcal_consumed, reach_meal_water_grams_consumed, reach_meal_water_kcal_consumed, cebq_sr, cebq_avoid, cebq_eue, cebq_ff, cebq_fr, cebq_se, meal_grams_consumed_inc_water, eah_grams_consumed_inc_water,total_grams_consumed_inc_water, total_kcal_consumed, reach_meal_tender_grams_consumed, reach_meal_tender_kcal_consumed, total_grams_consumed)) %>% tbl_summary()
```

6.  Generating demographics table by `study` with medians. `Unknown` indicates missing data

```{r}
thesis_data %>% select (!c(participant_id,pre_meal_fullness,pre_eah_fullness, meal_grams_consumed, meal_kcal_consumed, eah_grams_consumed_foodonly,bis, bas,bas_funseeking, bas_drive, bas_rewardresp, eah_kcal_consumed,fb_meal_consumed_applesauce_kcal, fb_meal_consumed_applesauce_g, fb_meal_consumed_carrot_g, fb_meal_consumed_carrot_kcal, fb_meal_consumed_cheese_sndwch_g, fb_meal_consumed_cheese_sndwch_kcal, fb_meal_consumed_cookies_g, fb_meal_consumed_cookies_kcal, fb_meal_consumed_ham_sndwch_g, fb_meal_consumed_ham_sndwch_kcal, fb_meal_consumed_milk_g, fb_meal_consumed_milk_kcal, fb_meal_consumed_pbj_sndwch_g, fb_meal_consumed_pbj_sndwch_kcal, fb_meal_consumed_potatochip_g, fb_meal_consumed_potatochip_kcal, fb_meal_consumed_turkey_sndwch_g, fb_meal_consumed_turkey_sndwch_kcal, fb_meal_consumed_ketchup_g, fb_meal_consumed_ketchup_kcal, fb_meal_consumed_mayo_g, fb_meal_consumed_mayo_kcal, fb_meal_consumed_mustard_g, fb_meal_consumed_mustard_kcal,fb_eah_consumed_brownies_g, fb_eah_consumed_brownies_kcal, fb_eah_consumed_cornchips_g,fb_eah_consumed_cornchips_kcal,fb_eah_consumed_hersheys_g,fb_eah_consumed_hersheys_kcal,fb_eah_consumed_icecream_g,fb_eah_consumed_icecream_kcal,fb_eah_consumed_oreos_g, fb_eah_consumed_oreos_kcal, fb_eah_consumed_popcorn_g,fb_eah_consumed_popcorn_kcal, fb_eah_consumed_pretzels_g, fb_eah_consumed_pretzels_kcal, fb_eah_consumed_skittles_g, fb_eah_consumed_skittles_kcal, fb_eah_consumed_starbursts_g,fb_eah_consumed_starbursts_kcal, fb_eah_consumed_water_g, reach_eah_brownie_grams_consumed, reach_eah_brownie_kcal_consumed, reach_eah_corn_chip_grams_consumed, reach_eah_corn_chip_kcal_consumed, reach_eah_ice_cream_grams_consumed,reach_eah_ice_cream_kcal_consumed,reach_eah_kiss_grams_consumed,reach_eah_kiss_kcal_consumed,reach_eah_oreo_grams_consumed, reach_eah_oreo_kcal_consumed, reach_eah_popcorn_grams_consumed, reach_eah_popcorn_kcal_consumed, reach_eah_pretzel_grams_consumed, reach_eah_pretzel_kcal_consumed, reach_eah_skittle_grams_consumed, reach_eah_skittle_kcal_consumed, reach_eah_starburst_grams_consumed, reach_eah_starburst_kcal_consumed,reach_eah_water_eah_grams_consumed,reach_eah_water_eah_kcal_consumed, reach_meal_grilled_cheese_grams_consumed,reach_meal_grilled_cheese_kcal_consumed, reach_meal_carrot_grams_consumed, reach_meal_carrot_kcal_consumed, reach_meal_chips_grams_consumed, reach_meal_chips_kcal_consumed, reach_meal_fruit_grams_consumed, reach_meal_fruit_kcal_consumed, reach_meal_ketchup_grams_consumed, reach_meal_ketchup_kcal_consumed, reach_meal_ranch_grams_consumed, reach_meal_ranch_kcal_consumed, reach_meal_water_grams_consumed, reach_meal_water_kcal_consumed, cebq_sr, cebq_avoid, cebq_eue, cebq_ff, cebq_fr, cebq_se, meal_grams_consumed_inc_water, eah_grams_consumed_inc_water,total_grams_consumed_inc_water, total_kcal_consumed, reach_meal_tender_grams_consumed, reach_meal_tender_kcal_consumed, total_grams_consumed)) %>% tbl_summary(by = Study)

```

7.  Generating demographics table by `study` with means and SD. `Unknown` indicates missing data

```{r}
thesis_data %>% 
  select(!c(participant_id, pre_meal_fullness, pre_eah_fullness, meal_grams_consumed, 
            meal_kcal_consumed, eah_grams_consumed_foodonly, bis, bas, bas_funseeking, 
            bas_drive, bas_rewardresp, eah_kcal_consumed, fb_meal_consumed_applesauce_kcal, 
            fb_meal_consumed_applesauce_g, fb_meal_consumed_carrot_g, 
            fb_meal_consumed_carrot_kcal, fb_meal_consumed_cheese_sndwch_g, 
            fb_meal_consumed_cheese_sndwch_kcal, fb_meal_consumed_cookies_g, 
            fb_meal_consumed_cookies_kcal, fb_meal_consumed_ham_sndwch_g, 
            fb_meal_consumed_ham_sndwch_kcal, fb_meal_consumed_milk_g, 
            fb_meal_consumed_milk_kcal, fb_meal_consumed_pbj_sndwch_g, 
            fb_meal_consumed_pbj_sndwch_kcal, fb_meal_consumed_potatochip_g, 
            fb_meal_consumed_potatochip_kcal, fb_meal_consumed_turkey_sndwch_g, 
            fb_meal_consumed_turkey_sndwch_kcal, fb_meal_consumed_ketchup_g, 
            fb_meal_consumed_ketchup_kcal, fb_meal_consumed_mayo_g, fb_meal_consumed_mayo_kcal, 
            fb_meal_consumed_mustard_g, fb_meal_consumed_mustard_kcal, 
            fb_eah_consumed_brownies_g, fb_eah_consumed_brownies_kcal, 
            fb_eah_consumed_cornchips_g, fb_eah_consumed_cornchips_kcal, 
            fb_eah_consumed_hersheys_g, fb_eah_consumed_hersheys_kcal, 
            fb_eah_consumed_icecream_g, fb_eah_consumed_icecream_kcal, 
            fb_eah_consumed_oreos_g, fb_eah_consumed_oreos_kcal, fb_eah_consumed_popcorn_g, 
            fb_eah_consumed_popcorn_kcal, fb_eah_consumed_pretzels_g, 
            fb_eah_consumed_pretzels_kcal, fb_eah_consumed_skittles_g, 
            fb_eah_consumed_skittles_kcal, fb_eah_consumed_starbursts_g, 
            fb_eah_consumed_starbursts_kcal, fb_eah_consumed_water_g, 
            reach_eah_brownie_grams_consumed, reach_eah_brownie_kcal_consumed, 
            reach_eah_corn_chip_grams_consumed, reach_eah_corn_chip_kcal_consumed, 
            reach_eah_ice_cream_grams_consumed, reach_eah_ice_cream_kcal_consumed, 
            reach_eah_kiss_grams_consumed, reach_eah_kiss_kcal_consumed, 
            reach_eah_oreo_grams_consumed, reach_eah_oreo_kcal_consumed, 
            reach_eah_popcorn_grams_consumed, reach_eah_popcorn_kcal_consumed, 
            reach_eah_pretzel_grams_consumed, reach_eah_pretzel_kcal_consumed, 
            reach_eah_skittle_grams_consumed, reach_eah_skittle_kcal_consumed, 
            reach_eah_starburst_grams_consumed, reach_eah_starburst_kcal_consumed, 
            reach_eah_water_eah_grams_consumed, reach_eah_water_eah_kcal_consumed, 
            reach_meal_grilled_cheese_grams_consumed, reach_meal_grilled_cheese_kcal_consumed, 
            reach_meal_carrot_grams_consumed, reach_meal_carrot_kcal_consumed, 
            reach_meal_chips_grams_consumed, reach_meal_chips_kcal_consumed, 
            reach_meal_fruit_grams_consumed, reach_meal_fruit_kcal_consumed, 
            reach_meal_ketchup_grams_consumed, reach_meal_ketchup_kcal_consumed, 
            reach_meal_ranch_grams_consumed, reach_meal_ranch_kcal_consumed, 
            reach_meal_water_grams_consumed, reach_meal_water_kcal_consumed, 
            cebq_sr, cebq_avoid, cebq_eue, cebq_ff, cebq_fr, cebq_se, 
            meal_grams_consumed_inc_water, eah_grams_consumed_inc_water, 
            total_grams_consumed_inc_water, total_kcal_consumed, reach_meal_tender_grams_consumed, 
            reach_meal_tender_kcal_consumed, total_grams_consumed)) %>% 
  tbl_summary(by = Study, 
              statistic = list(all_continuous() ~ "{mean} ({sd})", 
                               all_categorical() ~ "{n} / {N} ({p}%)"), 
              digits = all_continuous() ~ 2) %>% 
  add_overall() %>% 
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Participant's demographics by study**")

```
